Meeting --- Summary LLM

### 2024-05-13 meetings/xx-xx-2024.qmd
runs

- Gradient spikes first, then large step in decoder / encoder
- These steps seem to affect one hidden unit more strongly than the other
- Increase of Decoder Weight gradients seems to precede the increase of the -LL
- The other gradients seem to react to this jump
- Gradient-Decoder $\to$ Step $\to$ Other gradients $\to$ Step ...

@@

>>>>

- Gradient spikes first, then large step in decoder / encoder
- These steps seem to affect one hidden unit more strongly than the other
- Increase of Decoder Weight gradients seems to precede the increase of the -LL
- The other gradients seem to react to this jump
- Gradient-Decoder $\to$ Step $\to$ Other gradients $\to$ Step ...
- [`torch.nn.utils.clip_grad_norm_(ae.parameters(), max_norm=dMaxGrad)`](https://pytorch.org/docs/stable/generated/torch.nn.utils.clip_grad_norm_.html#torch.nn.utils.clip_grad_norm_) seems to help. This clips the gradient if the norm exceeds `dMaxGrad`. [See here](https://www.cs.toronto.edu/~rgrosse/courses/csc321_2017/readings/L15%20Exploding%20and%20Vanishing%20Gradients.pdf#page=6.78).

@@

![Fixed clipping at 0.3. Later in the optimization this is too large.](./figures/dashboard_clipping_fixed.png)

Likely, you only want to clip true "outliers" in the gradients and not be too conservative.

![You can overdo it with the gradient clipping.](./figures/dashboard_clip_01.png)
